---
title: "Water Quality Analysis"
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
---

Downloaded Hancock Biological Station data for Kentucky Lake from https://www.murraystate.edu/qacd/cos/hbs/wq.cfm
Includes data for Water, Temp = "Temperature(Deg C)", SpC = "Specific Conductance(uS/m)", DO = "Dissolved Oxygen(mg/L)", pH = "pH(SU)", ORP = "Oxidation-Reduction Potential(mV)", Turb = "Turbidity(NTU)" from top and bottom of lake.

```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```

### Useful packages

```{r import packages, results='hide'}
if (!require("pacman"))
   install.packages("pacman")

pacman::p_load(tidyverse, EnvStats, GGally, skimr, lubridate, forcats, timetk, mblm)
```

### Read data files
Downloaded data from https://www.murraystate.edu/qacd/cos/hbs/wq.cfm

```{r input}
Surface <- read_csv("CRROutputSurface.csv", 
    col_types = cols(Date = col_date(format = "%m/%d/%y")))
Bottom <- read_csv("CRROutputBottom.csv", 
    col_types = cols(Date = col_date(format = "%m/%d/%y")))
```

```{r}
Surface <- select(Surface, -"Chlorophyll a (ug/L)")
Surface$Loc <- c("Surface")
Bottom$Loc <- c("Bottom")
Water <- bind_rows(Surface, Bottom)
Water <-  dplyr::rename(Water, Temp = "Temperature(Deg C)", SpC = "Specific Conductance(uS/m)", DO = "Dissolved Oxygen(mg/L)", pH = "pH(SU)", ORP = "Oxidation-Reduction Potential(mV)", Turb = "Turbidity(NTU)")
skim(Water)
Water[Water<=0] <- NA
Water2<-Water[complete.cases(Water),]
Water2$logTurb <- log10(Water2$Turb)
Water2 <- select(Water2, -("Turb"))
skim(Water2)
Water2
```

## Boxplots comparing surface and bottom

```{r}
SpC <- match("SpC", names(Water2))
pH <- match("pH", names(Water2))
ORP <- match("ORP", names(Water2))
DO <-  match("DO", names(Water2))
Temp <-  match("Temp", names(Water2))
logTurb <-  match("logTurb", names(Water2))
vars <- c(SpC, pH, ORP, DO, Temp, logTurb)
for (i in vars) {
   print(ggplot(Water2, aes_(x = ~Loc, y = as.name(names(Water2)[i]), color = ~Loc)) +
    geom_boxplot() + 
    stat_n_text())
 }
```

Not much difference in water quality between top and bottom of lake. That's because SURFACE at 2.231 meters	and BOTTOM at 2.902 meters; both in the epilimnion. Station location not given on website, but I assume it is next to the station. The lake must be shallow there.

### Correlogram

```{r Correlogram}
textvars2 <- c("SpC", "pH", "Temp")
ggpairs(Water2, columns = textvars2, aes(color=Loc))
```

### Temporal trends

```{r time trends}
vars2 <- c(SpC, pH, Temp)
for (i in vars2) {
   print(ggplot(Water2, aes_(x = ~Date, y = as.name(names(Water2)[i]), color = ~Loc, shape = ~Loc)) +
    geom_point()
   )
}
```

### Extract surface water data

```{r}
Surface2 <- filter(Water2, Loc == "Surface")
Surface2 <- select(Surface2, -c("Loc", "logTurb", "ORP", "DO"))
# Calculate # of years elapsed since beginning of measurements
Surface2$Years <- as.numeric((Surface2$Date - Surface2$Date[1])/365)
# skim(Surface2)
```

### Summary Tables for surface water

```{r summary tables}
textvars <- c("SpC", "pH", "Temp")
summary(Surface2[textvars])
```

### Time series analysis of surface water

```{r Seasonal decomposition of conductivity}
KYlakeSpC <- ts(Surface2$SpC, start = c(2005), frequency = 365)
plot(KYlakeSpC)
fit <- stl(KYlakeSpC, s.window = "periodic")
plot(fit)
```

Strong seasonal signal in SpC. Time series show this is also true for temperature.

### Use timetk package to analyze T, SpC and pH

```{r surface seasonal diagnostics}
# See https://www.rdocumentation.org/packages/timetk/versions/2.6.1
df_timetk <-  gather(Surface2, variable, measurement, Temp:pH, factor_key=TRUE)
df_timetk %>%
  group_by(variable) %>%
  plot_time_series(Date, measurement, 
                   .facet_ncol = 2, .facet_scales = "free",
                   .interactive = TRUE)

```

### Conductivity seasonal diagnostics

```{r conductivity seasonal diagnostics}
plot_seasonal_diagnostics(Surface2, Date, SpC)
```

### Temp seasonal diagnostics

```{r Temp seasonal diagnostics}
plot_seasonal_diagnostics(Surface2, Date, Temp)
```

### pH seasonal diagnostics

```{r pH seasonal diagnostics}
plot_seasonal_diagnostics(Surface2, Date, pH)
```

### Use Thiel-Sen regression to estimate rate of change in parameter equal to the slope of the parameter across time (Kaushal et al., 2018)

```{r SpC Thiel-Sen regression}
# mblm(SpC~Years, Surface2)
```
Thiel-Sen slope is negative, meaning conductivity has decreased over time.

```{r Temp Thiel-Sen regression}
# mblm(Temp~Years, Surface2)
```

Temperature has also decreased.

```{r pH Thiel-Sen regression}
# mblm(pH~Years, Surface2)
```

pH also decreased. So none of the trends agree with those inferred on a national scale by Kaushal et al. (2018).

```{r}
sessionInfo()
save.image()
```
