---
title: "HRP Stream Sample Analysis 2020-2021"
author: "Hannah Zanibi, VU Class of 2021"
output:
  html_document:
    toc: yes
    highlight: tango
  pdf_document:
    toc: yes
Data File: https://vanderbilt365-my.sharepoint.com/:x:/g/personal/hannah_r_zanibi_vanderbilt_edu/Edq4CKkTZGtMtO1K-u8P27QBIIH_k9BSQfJ_5gO187Qplw?e=B6NaCI
---

This notebook displays script and analytical results of 12 stream samples taken during a 44-hour sampling campaign from October 2-4, 2020. This campaign look place along Big East Fork Rd. in the Harpeth River Watershed, TN. The variations in stream hydrochemistry are displayed in two forms: 

1. Time-series mosaic
2. Indexed chart 

>The selected parameters include: nutrient concentrations, log pCO2 (bar), pH, d13C, as well as calcite and dolomite saturation indices. 

```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(fig.path = "Users/hanzanibi/HRP/Outputs")
```

```{r setup, echo=FALSE, include=FALSE}
library(tidyverse, warn.conflicts = FALSE)
library(lubridate)
library(ggplot2)
library(readxl)
library(rcartocolor)
library(knitr)
library(cowplot)

```
Setting up Color Palette
---

The plots created in this notebook use a colorblind-friendly color palette coded below.
```{r}
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499",
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888", "#0072B2", "#D55E00")
scales::show_col(safe_colorblind_palette)
```


Importing Stream Data
---
```{r}
WaterData <- read_excel("202010CSSWaterData_04.xlsx")
head(WaterData)
```
Output Table
---
```{r}
HRPOutputTable <- select(WaterData, SampleID, Date, Time, pH, logpCO2, 18:ncol(WaterData))
knitr::kable(HRPOutputTable)
write.csv(HRPOutputTable, file = "HRPOutputTable.csv")
```

Variable Selection
---
```{r}
WaterVar <- select(WaterData, SampleID,Time, d13C, pH, Temperature, logpCO2, SIcalcite, SIdolomite)
WaterVarPlot <- WaterVar %>%
  gather('pH','d13C','logpCO2','SIcalcite', 'SIdolomite','Temperature', key = "Parameter", value = "Concentration") 
WaterpHPlot <- WaterVar %>%
  gather('pH', 'Temperature', key = "Parameter", value = "Value") 
WaterTracePlot <- WaterData %>%
  gather( 'Se', key = "Parameter", value = "Concentration")
WaterPlotCombined <- WaterData %>%
  gather ('pH', 'logpCO2', 'SIcalcite', 'SIdolomite', 'Temperature', 'd13C', 'Cl', 'DIC', 'S', 'Se', 'Ca', 'Fe', 'Mg', 'NO3', 'P', 'K', key="Parameter", value="Value")
```

Trace Metal Plot
---
```{r}
Trace <- ggplot(WaterTracePlot, aes(x=fct_inorder(Time), y = Concentration, group=Parameter)) +
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_line(aes(y=Concentration), color="#CC6677", size=1, na.rm =TRUE) + 
  geom_point(shape=1,color="#CC6677", size=2, na.rm =TRUE) +
  ylab("Concentration ug/L") + 
  xlab("Time") + 
  facet_wrap(~Parameter, scale="free_y") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) + 
  geom_hline(yintercept=3.1, colour="black", size=1)+
  geom_text(aes(x=5,y=3.2),colour="black", size=3, label = "U.S. EPA Selenium Criterion for Protection of Aquatic Life")
Trace + theme(legend.position="none")
print(Trace)
```

Temperature and pH Plots
---
```{r}
pH <- ggplot(WaterpHPlot, aes(x=fct_inorder(Time), y = Value, color = Parameter, group=Parameter)) +
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_line(aes(y=Value, colour=Parameter), size=1, na.rm =TRUE) + 
  geom_point(aes(shape=Parameter), size=2, na.rm =TRUE) +
  ylab("Value") + 
  xlab("Time") + 
  facet_wrap(~Parameter, scale="free_y") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) + scale_color_manual(values=safe_colorblind_palette) + 
  scale_shape_manual(values=c(0, 1, 2, 5, 6))
pH + theme(legend.position="none")
print(pH)
```
```{r}
pHTemp <- ggplot(WaterVar, aes(x=Temperature, y=pH)) +
  geom_point() +
  geom_smooth(method = "lm", colour="#AA4499") +
  ylab("pH") + 
  xlab("Temperature (Deg C)") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) + 
  scale_color_manual(values=safe_colorblind_palette)
print(pHTemp)
```

Create Geochemical Parameter Time-Series Plots
---
>Make sure to include fct_inorder to ensure that the time-series is in the correct order

```{r}
#Add standard deviation data for d13C
d13CSD <- unlist(select(WaterData,d13CStDev))

p <- ggplot(WaterVarPlot, aes(x=fct_inorder(Time), y = Concentration, color = Parameter, group=Parameter)) +
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_line(aes(y=Concentration, colour=Parameter), size=1, na.rm =TRUE) + 
  geom_point(aes(shape=Parameter), size=2, na.rm =TRUE) +
  ylab("Value") + 
  xlab("Time") + 
  facet_wrap(~Parameter, scale="free_y") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 2)) +
  theme(text=element_text(family="Times New Roman", size=14)) + 
  geom_hline(data = subset(WaterVarPlot, Parameter == "logpCO2"), aes(yintercept = -3.3979758),linetype = "dotted") +
  geom_errorbar(data=subset(WaterVarPlot, Parameter == "d13C"), aes(ymin=Concentration-d13CSD, ymax=Concentration+d13CSD), width=.2, position=position_dodge(.9)) +
  geom_text(data = subset(WaterVarPlot, Parameter == "logpCO2"), aes(x=5,y=-3.3),colour="black", size=3, label = "Atmospheric pCO2") + 
  scale_color_manual(values=c("#CC6677", "#DDCC77","#88CCEE", "#117733", "#332288", "#AA4499")) + 
  scale_shape_manual(values=c(0, 1, 2, 3, 5, 6)) 
p + theme(legend.position="none")
print(p)
```

>Note that geom_hline is used to add a reference line for atmospheric pCO2 at -3.39...

Combined Parameter Plot
---
```{r}
d13CSD <- unlist(select(WaterData,d13CStDev))

All <- ggplot(WaterPlotCombined, aes(x=fct_inorder(Time), y = Value, group=Parameter)) +
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_line(aes(y=Value), size=1, color="black", na.rm =TRUE) + 
  geom_point(shape=1,size=2, na.rm =TRUE, color="black") +
  ylab("Value") + 
  xlab("Time") + 
  facet_wrap(~Parameter, scale="free_y") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 2)) +
  theme(text=element_text(family="Times New Roman", size=14)) + 
  geom_hline(data = subset(WaterPlotCombined, Parameter == "logpCO2"), aes(yintercept = -3.3979758),linetype = "dotted") +
  geom_errorbar(data=subset(WaterPlotCombined, Parameter == "d13C"), aes(ymin=Value-d13CSD, ymax=Value+d13CSD), width=.2, position=position_dodge(.9)) +
  geom_text(data = subset(WaterPlotCombined, Parameter == "logpCO2"), aes(x=5,y=-3.3),colour="black", size=3, label = "Atmospheric pCO2") 
All + theme(legend.position="none")
print(All)
```
Create Index for Selected Parameters
---
```{r}
ParameterInd <- WaterVarPlot %>% 
  group_by(Parameter) %>% 
  mutate(ind = Concentration / mean(Concentration))
PercentParInd <- ParameterInd %>%
  mutate(perind= (ind-1)*100)
```
This will use the same parameters (pH, d13C, Dolomite, Calcite, and logpCO2) as before 

Create Geochemical Parameter Index Chart 
---

```{r}
p2 <- ggplot(ParameterInd, aes(x=fct_inorder(Time), y = (ind-1)*100, colour=Parameter, group=Parameter)) + 
  theme_bw() +
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.006) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.006) +
  geom_line(size=1) +
  geom_point(aes(shape=Parameter), size=2) +
  ylab("% Change from Mean") + 
  xlab("Time") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) + 
  scale_color_manual(values=c("#DDCC77","#88CCEE", "#117733", "#332288", "#AA4499","#882255")) + 
  scale_shape_manual(values=c(0, 1, 2, 3, 5, 6))
print(p2)
```
Element Selection 
---
Elements that displayed a (at minimum) weak diurnal over time were selected, with the exception of the nutrient parameters, which were selected to analyze nutrient variation at the stream site. 

The plots will be divided into anion concentrations (S, Se, Cl, DIC), cation concentrations (Ca, Fe, Mg), and nutrient concentrations (K, P, NO3).
```{r}
WaterNutrient <- select(WaterData, Time, K, P, NO3, Ca, Fe, Mg, S, Se, Cl, DIC, pH, logpCO2, Temperature, SIcalcite, SIdolomite, hydroxyapatite)
WaterNutrientPlot1 <- WaterNutrient %>%
  gather('S','Se', 'Cl', 'DIC', key = "Anion", value = "Concentration") 
WaterNutrientPlot2 <- WaterNutrient %>%
  gather('Ca','Fe', 'Mg', key = "Cation", value = "Concentration") 
WaterNutrientPlot3 <- WaterNutrient %>%
  gather('K', 'P', 'NO3', key = "Nutrient", value = "Concentration") 
#Nutrient plot w/o K 
WaterNutrientPlot4 <- WaterNutrient %>%
  gather('P', 'NO3', key = "Nutrient", value = "Concentration") 
```
Time-Series Plots for Element Data
---
```{r}
p3 <- ggplot(WaterNutrientPlot1, aes(x=fct_inorder(Time), y = Concentration, color = Anion, group=Anion)) +
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_line(aes(y=Concentration, colour=Anion), size=1) + 
  geom_point(aes(shape=Anion), size=2) +
  ylab("Concentration (ug/L)") + 
  xlab("Time") + 
  facet_wrap(~Anion, scale="free_y") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) +
  scale_color_manual(values=safe_colorblind_palette) +
  scale_shape_manual(values=c(0, 1, 2, 5, 6))
p3 + theme(legend.position="none")
print(p3)
```
```{r}
p4 <- ggplot(WaterNutrientPlot2, aes(x=fct_inorder(Time), y = Concentration, color = Cation, group=Cation)) +
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_line(aes(y=Concentration, colour=Cation), size=1) + 
  geom_point(aes(shape=Cation), size=2) +
  ylab("Concentration (ug/L)") + 
  xlab("Time") + 
  facet_wrap(~Cation, scale="free_y") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) + 
  scale_color_manual(values=safe_colorblind_palette) +
  scale_shape_manual(values=c(0, 1, 2, 5, 6))
p4 + theme(legend.position="none")
print(p4)
```
```{r}
p5 <- ggplot(WaterNutrientPlot3, aes(x=fct_inorder(Time), y = Concentration, color = Nutrient, group=Nutrient)) +
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.05) +
  geom_line(aes(y=Concentration, colour=Nutrient), size=1) + 
  geom_point(aes(shape=Nutrient), size=2) +
  ylab("Concentration (ug/L)") + 
  xlab("Time") + 
  facet_wrap(~Nutrient, scale="free_y") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) + 
  scale_color_manual(values=safe_colorblind_palette) +
  scale_shape_manual(values=c(0, 1, 2, 5, 6))
p5 + theme(legend.position="none")
print(p5)
```

Create Index for Element Data
---
```{r}
NutrientInd1 <-  
  WaterNutrientPlot1 %>% 
  group_by(Anion) %>% 
  mutate(ind = Concentration / mean(Concentration))
PercentAnInd <- NutrientInd1 %>%
  mutate(perind= (ind-1)*100)
NutrientInd2 <-  
  WaterNutrientPlot2 %>% 
  group_by(Cation) %>% 
  mutate(ind = Concentration / mean(Concentration))
PercentCatInd <- NutrientInd2 %>%
  mutate(perind= (ind-1)*100)
NutrientInd3 <-  
  WaterNutrientPlot3 %>% 
  group_by(Nutrient) %>% 
  mutate(ind = Concentration / mean(Concentration))
NutrientInd4 <-  
  WaterNutrientPlot4 %>% 
  group_by(Nutrient) %>% 
  mutate(ind = Concentration / mean(Concentration))
```

Create Nutrient Index Charts
---
```{r}
p6 <- ggplot(NutrientInd1, aes(x=fct_inorder(Time), y = (ind-1)*100, colour=Anion, group=Anion)) + 
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.01) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.01) +
  geom_line(size=1) +
  geom_point(aes(shape=Anion), size=2) + 
  ylab("% Change from Mean ") + 
  xlab("Time") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) +
  scale_color_manual(values=safe_colorblind_palette) +
  scale_shape_manual(values=c(0, 1, 2, 5, 6))
print(p6)
```

```{r}
p7 <- ggplot(NutrientInd2, aes(x=fct_inorder(Time), y = (ind-1)*100, colour=Cation, group=Cation)) + 
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.01) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.01) +
  geom_line(size=1) +
  geom_point(aes(shape=Cation), size=2) + 
  ylab("% Change from Mean ") + 
  xlab("Time") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) + 
  scale_color_manual(values=safe_colorblind_palette) +
  scale_shape_manual(values=c(0, 1, 2, 5, 6))
print(p7)
```
>Note: Because K does not exhibit a diurnal signal (shown in plot 5), K is removed from the 'Relative Change in Nutrients' plot

```{r}
p8 <- ggplot(NutrientInd4, aes(x=fct_inorder(Time), y = (ind-1)*100, colour=Nutrient, group=Nutrient)) + 
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.01) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.01) +
  geom_line(size=1) +
  geom_point(aes(shape=Nutrient), size=2) + 
  ylab("% Change from Mean ") + 
  xlab("Time") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) + 
  scale_color_manual(values=safe_colorblind_palette) +
  scale_shape_manual(values=c(0, 1, 2, 5, 6))
print(p8)
```

Magnitude Change 
---
>The following formula is used to calculate the magnitude of diel variations as done in Nimick, 2011

```{r}
MagnChange1 <-  
  WaterVarPlot %>% 
  group_by(Parameter) %>% 
  mutate(ind = ((max(Concentration)-min(Concentration))/min(Concentration)*100))
MagnChange2 <-  
  WaterpHPlot %>% 
  group_by(Parameter) %>% 
  mutate(ind = ((max(Value)-min(Value))/min(Value)*100))
MagnChange3 <-  
  WaterNutrientPlot1 %>% 
  group_by(Anion) %>% 
  mutate(ind = ((max(Concentration)-min(Concentration))/min(Concentration)*100))
MagnChange4 <-  
  WaterNutrientPlot2 %>% 
  group_by(Cation) %>% 
  mutate(ind = ((max(Concentration)-min(Concentration))/min(Concentration)*100))
MagnChange5 <-  
  WaterNutrientPlot3 %>% 
  group_by(Nutrient) %>% 
  mutate(ind = ((max(Concentration)-min(Concentration))/min(Concentration)*100))
MagnChangeCombined <-
  WaterPlotCombined %>%
  group_by(Parameter)%>%
  mutate(ind = ((max(Value)-min(Value))/min(Value)*100))
```

Selenium Calculations
---
```{r}
SeleniumPlot <- WaterNutrient %>%
  gather('Fe', 'pH', key = "Parameter", value = "Value") 
SeleniumInd <-  
  SeleniumPlot %>% 
  group_by(Parameter) %>% 
  mutate(ind = Value / mean(Value))
p9 <- ggplot(SeleniumInd, aes(x=fct_inorder(Time), y = (ind-1)*100, colour=Parameter, group=Parameter)) + 
  geom_rect(aes(xmin='17:50', xmax='05:58', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.01) +
  geom_rect(aes(xmin='18:09', xmax='05:52', ymin=-Inf, ymax=Inf), fill='grey', colour=NA, alpha=0.01) +
  geom_line(size=1) +
  geom_point(aes(shape=Parameter), size=2) + 
  ylab("% Change from Mean ") + 
  xlab("Time") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(text=element_text(family="Times New Roman", size=14)) + 
  scale_color_manual(values=safe_colorblind_palette) +
  scale_shape_manual(values=c(0, 1, 2, 5, 6))

print(p9)

```

Correlation Calculations 
---
```{r}
library(corrplot)
SeleniumPlot2 <- WaterNutrient %>%
  select('K', 'P', 'NO3', 'Ca', 'Fe', 'Mg', 'S', 'Se', 'Cl', 'DIC', 'pH', 'logpCO2', 'Temperature', 'SIcalcite', 'SIdolomite', 'hydroxyapatite')
SeCorr<-cor(SeleniumPlot2)
SeCorr
head(round(SeleniumPlot2,2))
corrplot(SeCorr, method="circle")

```

```{r}
cex.before <- par("cex")
par(cex = 0.6)
p10 <- corrplot(SeCorr, insig = "blank", method = "color", col=colorRampPalette(c("#D55E00","white","#0072B2"))(200),
    addCoef.col="black", tl.col="black",
    order = "AOE", tl.cex = 1/par("cex"),
    cl.cex = 1/par("cex"),  addCoefasPercent = FALSE)
dev.copy(svg, "CorrelationMatrix.svg")
par(cex = cex.before)
```
> Methods obtained from http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram 
 
Display Session Info 
---
```{r}
```


```{r}
sessionInfo()
```

