---
title: "Mill Creek Stream Survey"
output:
  html_document:
    df_print: paged
---

Creating maps using methods described in Brunsdon and Comer (2019).

```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```

### Set working directory and load packages

```{r}

setwd("C:/Users/ayersj.VANDERBILT/OneDrive - Vanderbilt/Shared Documents/Pilot Project - Water Quality/Data/2022/Survey")
if (!require("pacman"))
  install.packages("pacman")
pacman::p_load(tidyverse, readxl, sf, sp, rgdal, tmap, GISTools, RgoogleMaps, leaflet, skimr, raster, corrplot)

```

### Import and summarize data

```{r}
Survey <- read_excel("MC_Survey_V2.xlsx", 
     col_types = c("date", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric"))
Survey <- dplyr::select(Survey, -ConduSPercm, -TDSmgPerL)
skimr::skim(Survey)

```

### Use tmap to create a map
Lose most map elements when I add background using leaflet

```{r}
# define the coordinates
coords.tmp <- cbind(Survey$Longitude, Survey$Latitude)
# create the SpatialPointsDataFrame
Stream.sp <- SpatialPointsDataFrame(coords.tmp, 
                                    data = data.frame(Survey), 
                                    proj4string = CRS("+proj=longlat"))
# convert to sf
Stream_sf <- st_as_sf(Stream.sp)
# create map
tmap_mode('view') # Launches leaflet, but then lose most customizations
# Saving as a png without leaflet loses most elements
# png("SurveyTmapLayout.png", width = 2000, height = 2*618, units = "px")
tm_shape(Stream_sf) +
  tm_dots("ODOmgPerL", shape = 19, alpha = 0.7, size = 0.1, 
          palette = "Blues", 
          title="Dissolved O2 (mg/L)",
          style = "kmeans", legend.hist = T)  +
  tm_scale_bar(width = 0.22) +
  tm_compass(position = c(0.8, 0.07)) +
  tm_layout(frame = F, title = "Mill Creek", 
            title.size = 2, title.position = c(0.55, "top"),
            legend.width = 0.5,
            legend.hist.size = 0.5)
# dev.off()
# tmap_mode("plot")

```

### Create map using RgoogleMaps package. 
Don't like because it uses base R syntax.

```{r}
# define Lat and Lon
Lat <- as.vector(Survey$Latitude)
Long <- as.vector(Survey$Longitude)
DO <- Survey$ODOmgPerL
# From https://r-graph-gallery.com/13-scatter-plot.html
# Create a color palette
mypalette <- colorNumeric( palette="YlOrBr", domain=Survey$ODOmgPerL, na.color="transparent")

# get the map tiles 
MyMap <- MapBackground(lat=Lat, lon=Long)
PlotOnStaticMap(MyMap, Lat, Long, cex=DO/10, pch=16, col=mypalette(Survey$ODOmgPerL))

```

### Create leaflet map

```{r}
# See https://r-graph-gallery.com/19-map-leafletr.html
# Create a color palette with handmade bins.
mybins <- seq(8, 16, by=2)
mypalette <- colorBin( palette="YlOrBr", domain=Survey$ODOmgPerL, na.color="transparent", bins=mybins)
# Prepare the text for the tooltip:
mytext <- paste(
   "DO % Sat: ", Survey$ODOPercSat, "<br/>", 
   "ORP_mV: ", Survey$ORP_mV, "<br/>", 
   "pH: ", Survey$pH, sep="") %>%
  lapply(htmltools::HTML)
# Final Map
m <- leaflet(Survey) %>% 
  addTiles()  %>% 
  setView( lat=36.1, lng=-86.7, zoom=12) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addCircleMarkers(~Longitude, ~Latitude, 
    fillColor = ~mypalette(ODOmgPerL), fillOpacity = 0.7, color="white", radius=8, stroke=FALSE, label = mytext,
    labelOptions = labelOptions( style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "13px", direction = "auto")
  ) %>%
  addLegend( pal=mypalette, values=~ODOmgPerL, opacity=0.9, title = "DO (mg/L)", position = "bottomright" )
m 

```

### Calculate the cumulative distance of each point along the stream from the first point.

```{r}
is.projected(Stream.sp) # Returns FALSE, so must project to UTM Zone 16S as determined from https://www.latlong.net/lat-long-utm.html
# EPSG = 32716 https://epsg.io/32716
# Calculate distances, first using sf
raster::extent(Stream.sp)
crs(Stream.sp)
# proj=longlat: the data are in a geographic (latitude and longitude) coordinate system
Dist <- st_distance(Stream_sf, which = ifelse(isTRUE(st_is_longlat(Stream_sf)), "Great Circle", "Euclidean"))
#Distance from first to last point is 6170m
# The following saves the distance of each point from the first point
# Survey$Dist_m <- as.numeric(Dist[,1])
#Actually want the distance from each point to the next point (from 1-2, 2-3, etc.). 
l = nrow(Dist)-1
ConsecDist = c(0)
for (i in 1:l) {
  ConsecDist[i+1] = Dist[i+1,i]
}
Survey$Dist_m <- ConsecDist
Survey$CumDist <- cumsum(Survey$Dist_m)
# See https://www.earthdatascience.org/courses/earth-analytics/spatial-data-r/reproject-vector-data/
# UTM <- CRS("+proj=utm +zone=16 +south +datum=WGS84 +units=m +no_defs ")
# Stream.sp_UTM <- spTransform(Stream.sp, UTM)
# is.projected(Stream.sp_UTM)
# crs(Stream.sp_UTM)
# gDistance(Stream.sp_UTM)
# as.matrix(dist(coordinates(Stream.sp_UTM)))
#Distance from first to last point is 6161m. Close enough.

```

### Make distance plots

```{r, error=FALSE, warning=FALSE}
SurveyLong <- Survey %>%
  dplyr::select(ODOPercSat:Temp_C, Altitude_m, CumDist) %>%
  gather(ODOPercSat:Altitude_m, key = "Parameter", value = "Value")
print(ggplot(SurveyLong, aes(x = CumDist/1000, y = Value)) +
      geom_point(na.rm = TRUE) +
      geom_line() +
      facet_wrap(facets = vars(Parameter), scales = "free_y") +
      xlab("Distance along stream (km)") +
      theme(axis.text.x = element_text(angle = 90))
      )

```

### Create correlation matrix

```{r}
Parameters <- dplyr::select(Survey, ODOPercSat:Temp_C)
Corr <- cor(Parameters, use = "pairwise.complete.obs", method = "pearson")
corrplot(Corr, method="circle")
```


### Save output file and display session info

```{r}
save.image("202204_MC_Survey.RData")
sessionInfo()
```
