---
title: "Mill Creek Analysis of Seasonality and Diel Variation of Water Quality Parameters 2019-2020"
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
---

This notebook investigates the seasonality of four water quality measures that have been measured by the USGS since 2007: temperature (degrees Celsius), conductivity (a measure of salinity in microsiemens per centimeter), pH, and dissolved oxygen (mg/L). The site is USGS 03431083 MILL CREEK AT LEBANON ROAD AT NASHVILLE, TN. Information on the site is at <https://waterdata.usgs.gov/nwis/inventory/?site_no=03431083&agency_cd=USGS> Location: Latitude 36°09'50", Longitude 86°41'57" NAD27 This analysis is for winter 2019 through fall 2020.

```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE, echo=FALSE, message=FALSE, warning=FALSE)
```

```{r import packages, results='hide'}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, lubridate, skimr, timetk, leaflet, psych, htmltools)
```

### Map of site location with default Streetmap basemap

```{r Map}
site <- tibble(lat = 36.16389, lng = -86.69917, cap = c("USGS Station 03431083 Mill Creek at Lebanon Road, https://waterdata.usgs.gov/nwis/inventory/?site_no=03431083&agency_cd=USGS"))
m <- leaflet(site) %>% addMarkers(lat = ~lat, lng = ~lng, popup = ~htmlEscape(cap)) %>% addTiles()
m
```

### ESRI World basemap

```{r ESRI World basemap}
m %>% addProviderTiles(providers$Esri.WorldImagery)
```

```{r Read data from USGS website}
# Unfortunately this USGS station does not have discharge data
# The begin and end dates can be changed in the url below
Data <- read_delim("https://nwis.waterdata.usgs.gov/usa/nwis/uv/?cb_00095=on&cb_00300=on&cb_00010=on&cb_00400=on&format=rdb&site_no=03431083&period=&begin_date=2019-12-22&end_date=2020-12-21", "\t", escape_double = FALSE, col_names = TRUE, locale = locale(tz = "America/Chicago"), trim_ws = TRUE, skip = 30)
Data <- Data[-1,-(1:2)]
Data <- rename(Data, Temp_C = `131093_00010`, SpC = `131094_00095`, pH = `131095_00400`, DO = `131096_00300`)
Data <- select(Data, datetime, Temp_C, SpC, pH, DO)
Data <- type_convert(Data)
skim(Data)
write_excel_csv(Data, "MillCreekWQ_Seasonality.csv", col_names = TRUE)
```

### Plots of temporal trends

```{r temporal trends}
DataLong <-  gather(Data, variable, measurement, Temp_C:DO, factor_key=TRUE)
MinDate <- as.Date(min(Data$datetime))
MaxDate <- as.Date(max(Data$datetime))
print(paste("From ", MinDate, " to ", MaxDate))
print(ggplot(DataLong, aes(x = datetime, y = measurement)) +
  geom_line() +
  scale_x_datetime(date_labels = "%Y-%m") +
  facet_wrap(~ variable, scales = "free", nrow = 2))
```

-   see strong seasonality in temperature and dissolved oxygen. Conductivity/salinity shows negative spikes associated with precipitation events.

### Annual summary statistics

```{r Annual summary statistics}
textvars <- c("Temp_C", "SpC", "pH", "DO")
describeBy(Data[textvars])
```

```{r Create seasons}
Data <- Data %>%
  add_column(Season = NA)
for (i in 1:nrow(Data)) {
  if (as.Date(Data$datetime[i]) >= as.Date("2019-12-21") & as.Date(Data$datetime[i]) <= as.Date("2020-03-19")) {
  Data$Season[i] = paste("Winter")
  }
  if (as.Date(Data$datetime[i]) >= as.Date("2020-03-20") & as.Date(Data$datetime[i]) <= as.Date("2020-06-19")) {
  Data$Season[i] = paste("Spring")
  }
   if (as.Date(Data$datetime[i]) >= as.Date("2020-06-20") & as.Date(Data$datetime[i]) <= as.Date("2020-09-21")) {
  Data$Season[i] = paste("Summer")
  }
   if (as.Date(Data$datetime[i]) >= as.Date("2020-09-22") & as.Date(Data$datetime[i]) <= as.Date("2020-12-20")) {
  Data$Season[i] = paste("Fall")
   }
}
```

### Seasonal boxplots

```{r Seasonal boxplots}
Data$Season <- factor(Data$Season, levels = c("Spring", "Summer", "Fall", "Winter"))
Data2 <- Data %>% drop_na() 
SpC <- match("SpC", names(Data2))
pH <- match("pH", names(Data2))
DO <-  match("DO", names(Data2))
Temp_C <-  match("Temp_C", names(Data2))
vars <- c(Temp_C, DO, SpC, pH)
SeasonBoxplot <- function(z) {
  for (i in vars) {
   print(ggplot(z, aes_(x = ~Season, y = as.name(names(z)[i]), color = ~Season)) +
    geom_boxplot())
          }
}
SeasonBoxplot(Data2)
```

The width of the boxes in the boxplots above show how much a parameter changed within a season.

-   Temperature shows strong seasonality, being highest in the summer and lowest in the winter. It varies the most in the spring and fall.

-   Dissolved oxygen is the highest in the fall and winter when temperatures are the lowest and oxygen solubility is the highest. Stream is oxidizing, not eutrophic. DO varies the most in the spring and fall.

-   Conductivity/salinity is highest in fall when precipitation is lowest.

-   pH shows no seasonal variation.

## Daily variations in fall when diel cycles are strongest

### Temperature

```{r Temp hourly diagnostics}
Fall <- filter(Data, Season == "Fall")
plot_seasonal_diagnostics(Fall, datetime, Temp_C, .feature_set = c("hour"), .interactive = TRUE)
```

-   Results for temperature a little puzzling, with highest values in the evening through \~ 1 am, with the lowest value at 3 pm. This would probably be different if data for multiple years used; one-year dataset is too small.

### Dissolved oxygen

```{r DO daily diagnostics}
plot_seasonal_diagnostics(Fall, datetime, DO, .feature_set = c("hour"), .interactive = TRUE)
```

-   Dissolved oxygen reaches a minimum in the morning after being consumed at night by respiration, and the highest levels in the evening when the sun sets and photosynthesis ceases.

### Conductivity

```{r conductivity daily diagnostics}
plot_seasonal_diagnostics(Fall, datetime, SpC, .feature_set = c("hour"), .interactive = TRUE)
```

-   Conductivity shows very minor daily variation, with highest levels in the afternoon when evapotranspiration is highest.

### pH

```{r pH daily diagnostics}
plot_seasonal_diagnostics(Fall, datetime, pH, .feature_set = c("hour"), .interactive = TRUE)
```

-   clear diel signal for pH, with lowest values 6 am - 3 pm when photosynthesis consumes CO2 and highest after the sun sets from 8 pm - 2 am when respiration consumes CO2.

```{r}
# sessionInfo()
# save.image("MillCreekSeasonality.RData")
```
