---
title: "Mill Creek Water Analysis"
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
---

This notebook investigates four water quality measures that have been measured by the USGS since 2007: temperature, conductivity (a measure of salinity), pH, and dissolved oxygen. The site is USGS 03431083 MILL CREEK AT LEBANON ROAD AT NASHVILLE, TN. Information on the site is at https://waterdata.usgs.gov/nwis/inventory/?site_no=03431083&agency_cd=USGS
Location: Latitude 36°09'50",   Longitude 86°41'57"   NAD27

```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```

### Useful packages

```{r import packages, results='hide'}
if (!require("pacman"))
   install.packages("pacman")

pacman::p_load(tidyverse, skimr, timetk, leaflet)
```

### Map of site location

```{r Map}
m <- leaflet(df) %>% addMarkers(lat = 36.16389, lng = -86.69917) %>% addTiles()
m
```

### Import and clean data

```{r Read data from USGS website}
# Unfortunately this USGS station does not have discharge data
# The begin and end dates can be changed in the url below
Data <- read_delim("https://nwis.waterdata.usgs.gov/usa/nwis/uv/?cb_00095=on&cb_00300=on&cb_00010=on&cb_00400=on&format=rdb&site_no=03431083&period=&begin_date=2020-10-02&end_date=2020-10-04", "\t", escape_double = FALSE, col_names = TRUE, locale = locale(tz = "America/Chicago"), trim_ws = TRUE, skip = 30)
Data <- Data[-1,-(1:2)]
Data <- rename(Data, Temp_C = `131093_00010`, SpC = `131094_00095`, pH = `131095_00400`, DO = `131096_00300`)
Data <- select(Data, datetime, Temp_C, SpC, pH, DO)
Data <- type_convert(Data)
skim(Data)
# write_excel_csv(Data, "MillCreekDataWQ.csv", col_names = TRUE)
```

```{r temporal trends monthly timescale}
DataLong <-  gather(Data, variable, measurement, Temp_C:DO, factor_key=TRUE)
print(ggplot(DataLong, aes(x = datetime, y = measurement)) +
  geom_line() +
  scale_x_datetime(date_labels = "%Y-%m") +
  facet_wrap(~ variable, scales = "free", nrow = 2))
```

```{r temporal trends daily timescale}
# Set # of days to plot
ndays <- 3
# Measurements every 30 minutes, so for four days need last 4*48 = 192 rows
LastRows <- ndays*48
Week <- Data[-(1:(nrow(Data)-LastRows)),]
# Create index plot of % change from mean
percent_change <- function(x) {
  x <- 100*((x - mean(x, na.rm = TRUE))/mean(x, na.rm = TRUE))
}
WeekChange <- transmute(Week, datetime, 
                  Temp_C = percent_change(Temp_C),
                  SpC = percent_change(SpC),
                  pH = percent_change(pH),
                  DO = percent_change(DO)
                  )
WeekChange
```
### Summary Tables for surface water

```{r summary tables}
textvars <- c("Temp_C", "SpC", "pH", "DO")
summary(WeekChange[textvars])
```

```{r temporal trends daily timescale plot}
WeekLong <-  gather(WeekChange, variable, measurement, Temp_C:DO, factor_key=TRUE)
print(ggplot(WeekLong, aes(x = datetime, y = measurement, color = variable, shape = variable)) +
  geom_line() +
    ylab("% change")
)
```

### Conductivity seasonal diagnostics

```{r conductivity seasonal diagnostics}
plot_seasonal_diagnostics(Data, datetime, SpC, .feature_set = c("hour", "month.lbl"), .interactive = TRUE)
```

### Temp seasonal diagnostics

```{r Temp seasonal diagnostics}
plot_seasonal_diagnostics(Data, datetime, Temp_C, .feature_set = c("hour", "month.lbl"), .interactive = TRUE)
```

### pH seasonal diagnostics

```{r pH seasonal diagnostics}
plot_seasonal_diagnostics(Data, datetime, pH, .feature_set = c("hour", "month.lbl"), .interactive = TRUE)
```

### DO seasonal diagnostics

```{r DO seasonal diagnostics}
plot_seasonal_diagnostics(Data, datetime, DO, .feature_set = c("hour", "month.lbl"), .interactive = TRUE)
```

```{r}
sessionInfo()
save.image()
```