---
title: "July 2022 Time Series East Fork Creek"
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
---
This notebook analyzes chemical data from in-situ measurements and water samples collected from East Fork Creek in Franklin, TN by John C. Ayers, Weizhuo Jing, Kevin Chen, Ella Daugherty, and Gabriel Perez in July 2022. It shows the sample site in a map, and then creates time series plots illustrating the diel cycles. 

```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```

### Useful packages

```{r import packages, echo=FALSE, error=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, skimr, leaflet, scales, corrplot, car, timetk)

```

### Map of site location

```{r Map}
m <- leaflet(df) %>% addMarkers(lat = 35.951470, lng = -87.016110) %>% 
  addProviderTiles(providers$Esri.NatGeoWorldMap)
# See http://leaflet-extras.github.io/leaflet-providers/preview/index.html
m

```

```{r}
Data <- read_csv("EXO_Data_070622.csv", col_types = cols(Date = col_date(format = "%m/%d/%Y"), 
    Time = col_time(format = "%H:%M:%S"), 
    DateTime = col_datetime(format = "%m/%d/%Y %H:%M")), 
    locale = locale(tz = "America/Chicago"))
skim(Data)

```

### Grouped time series

```{r}
DataFacet <- transmute(Data, DateTime, days, Light_lux, Temp_C, pH, Depth_m, SpCond_uS_cm, ODO_PerLocSat, fDOM_RFU, Chlorophyll_RFU, TAL_PC_RFU,
  LightInt = rescale(Light_lux, to = c(0,1), from = range(Light_lux)),
  Light = if_else(Light_lux == 0, 1, 0)
  )
DataFacetLong <-  gather(DataFacet, variable, measurement, Light_lux:TAL_PC_RFU, factor_key=TRUE)
# Figure out how to shade using Light variable
p <- ggplot(DataFacetLong, aes(x=days)) +
  theme_linedraw() +
  geom_line(aes(y = measurement), color = "blue") +
  # geom_col(aes(y=Light), color = "grey", alpha = 0.1) +
  # scale_y_continuous(sec.axis = sec_axis(~.)) +
  facet_wrap(~ variable, scales = "free", nrow = 3)
print(p)
ggsave("2022_07_EFC_TimeSeries.png", path = "Plots")

```

### Temporal trends daily timescale

```{r}
percent_change <- function(x) {
  x <- 100*((x - mean(x, na.rm = TRUE))/mean(x, na.rm = TRUE))
}
PercChange <- transmute(DataFacet, days,
                  Light_lux = percent_change(Light_lux),
                  Temp_C = percent_change(Temp_C),
                  pH = percent_change(pH),
                  SpCond_uS_cm = percent_change(SpCond_uS_cm),
                  ODO_PerLocSat = percent_change(ODO_PerLocSat),
                  fDOM_RFU = percent_change(fDOM_RFU),
                  Chlorophyll_RFU = percent_change(Chlorophyll_RFU),
                  TAL_PC_RFU = percent_change(TAL_PC_RFU)
                  )
PercChangeLong <-  gather(PercChange, variable, measurement, Light_lux:TAL_PC_RFU)
p <- ggplot(PercChangeLong, aes(x = days, y = measurement, color = variable)) +
  theme_bw() +
    geom_line() +
    ylab("% change")
print(p)
ggsave("2022_07_EFC_TimeSeriesPercChange.png", path = "Plots")

```

### Correlation Matrix

```{r}
Parameters <- select(SelectData, -DateTime)
Corr <- cor(Parameters, use = "pairwise.complete.obs", method = "spearman")
corrplot(Corr, method="circle")
dev.copy(png, "Plots/CorrelationMatrix.png")
dev.off()

```

### Scatterplot Matrix

```{r}

scatterplotMatrix(~Light_lux + Temp_C + pH + Depth_m + SpCond_uS_cm + ODO_PerLocSat + fDOM_RFU + Chlorophyll_RFU + TAL_PC_RFU, data = Parameters)
dev.copy(png, "Plots/ScatterplotMatrix.png")
dev.off()

```

```{r}

# The following crams them all into one row; can't specify # of rows like facet_wrap
# DataFacetLong %>%
#   group_by(variable) %>%
#   plot_seasonal_diagnostics(DateTime, measurement, .feature_set = c("hour"), .interactive = FALSE, nrow = 3)

plot_seasonal_diagnostics(SelectData, DateTime, Light_lux, .feature_set = c("hour"), .interactive = TRUE, .title = "Light (lux)")
plot_seasonal_diagnostics(SelectData, DateTime, Temp_C, .feature_set = c("hour"), .interactive = TRUE, .title = "Temperature (°C)")
plot_seasonal_diagnostics(SelectData, DateTime, pH, .feature_set = c("hour"), .interactive = TRUE, .title = "pH")
plot_seasonal_diagnostics(SelectData, DateTime, SpCond_uS_cm, .feature_set = c("hour"), .interactive = TRUE, .title = "Specific Conductivity (μS/cm)")
plot_seasonal_diagnostics(SelectData, DateTime, ODO_PerLocSat, .feature_set = c("hour"), .interactive = TRUE, .title = "Dissolved Oxygen (% saturation)")
plot_seasonal_diagnostics(SelectData, DateTime, fDOM_RFU, .feature_set = c("hour"), .interactive = TRUE, .title = "Dissolved organic matter (RFU)")
plot_seasonal_diagnostics(SelectData, DateTime, Chlorophyll_RFU, .feature_set = c("hour"), .interactive = TRUE, .title = "Chlorophyll (RFU)")
plot_seasonal_diagnostics(SelectData, DateTime, TAL_PC_RFU, .feature_set = c("hour"), .interactive = TRUE, .title = "Total Algae (RFU)")

```


```{r}
sessionInfo()
save.image()

```

