---
title: "Mill Creek Analysis of Seasonality of Water Quality Parameters"
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
---

This notebook investigates the seasonality of four water quality measures that have been measured by the USGS since 2007: temperature (degrees Celsius), conductivity (a measure of salinity in microsiemens per centimeter), pH, and dissolved oxygen (mg/L). The site is USGS 03431083 MILL CREEK AT LEBANON ROAD AT NASHVILLE, TN. Information on the site is at https://waterdata.usgs.gov/nwis/inventory/?site_no=03431083&agency_cd=USGS
Location: Latitude 36°09'50",   Longitude 86°41'57"   NAD27

```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```

### Useful packages

```{r import packages, results='hide'}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, lubridate, skimr, timetk, leaflet, psych, htmltools)
```

### Map of site location with default Streetmap basemap

```{r Map}
site <- tibble(lat = 36.16389, lng = -86.69917, cap = c("USGS Station 03431083 Mill Creek at Lebanon Road, https://waterdata.usgs.gov/nwis/inventory/?site_no=03431083&agency_cd=USGS"))
m <- leaflet(site) %>% addMarkers(lat = ~lat, lng = ~lng, popup = ~htmlEscape(cap)) %>% addTiles()
m
```

### ESRI World basemap

```{r ESRI World basemap}
m %>% addProviderTiles(providers$Esri.WorldImagery)
```

### Import and clean data

```{r Read data from USGS website}
# Unfortunately this USGS station does not have discharge data
# The begin and end dates can be changed in the url below
Data <- read_delim("https://nwis.waterdata.usgs.gov/usa/nwis/uv/?cb_00095=on&cb_00300=on&cb_00010=on&cb_00400=on&format=rdb&site_no=03431083&period=&begin_date=2019-12-22&end_date=2020-12-21", "\t", escape_double = FALSE, col_names = TRUE, locale = locale(tz = "America/Chicago"), trim_ws = TRUE, skip = 30)
Data <- Data[-1,-(1:2)]
Data <- rename(Data, Temp_C = `131093_00010`, SpC = `131094_00095`, pH = `131095_00400`, DO = `131096_00300`)
Data <- select(Data, datetime, Temp_C, SpC, pH, DO)
Data <- type_convert(Data)
skim(Data)
write_excel_csv(Data, "MillCreekWQ_Seasonality.csv", col_names = TRUE)
```

### Plots of temporal trends

```{r temporal trends}
DataLong <-  gather(Data, variable, measurement, Temp_C:DO, factor_key=TRUE)
MinDate <- as.Date(min(Data$datetime))
MaxDate <- as.Date(max(Data$datetime))
print(paste("From ", MinDate, " to ", MaxDate))
print(ggplot(DataLong, aes(x = datetime, y = measurement)) +
  geom_line() +
  scale_x_datetime(date_labels = "%Y-%m") +
  facet_wrap(~ variable, scales = "free", nrow = 2))
```

### Create seasons

```{r Create seasons}
Data <- Data %>%
  add_column(Season = NA)
for (i in 1:nrow(Data)) {
  if (as.Date(Data$datetime[i]) >= as.Date("2019-12-21") & as.Date(Data$datetime[i]) <= as.Date("2020-03-19")) {
  Data$Season[i] = paste("Winter")
  }
  if (as.Date(Data$datetime[i]) >= as.Date("2020-03-20") & as.Date(Data$datetime[i]) <= as.Date("2020-06-19")) {
  Data$Season[i] = paste("Spring")
  }
   if (as.Date(Data$datetime[i]) >= as.Date("2020-06-20") & as.Date(Data$datetime[i]) <= as.Date("2020-09-21")) {
  Data$Season[i] = paste("Summer")
  }
   if (as.Date(Data$datetime[i]) >= as.Date("2020-09-22") & as.Date(Data$datetime[i]) <= as.Date("2020-12-20")) {
  Data$Season[i] = paste("Fall")
   }
}
```

### Annual summary statistics

```{r Annual summary statistics}
textvars <- c("Temp_C", "SpC", "pH", "DO")
describeBy(Data[textvars])
```

### Seasonal boxplots

```{r Seasonal boxplots}
Data$Season <- factor(Data$Season, levels = c("Spring", "Summer", "Fall", "Winter"))
Data2 <- Data %>% drop_na() 
SpC <- match("SpC", names(Data2))
pH <- match("pH", names(Data2))
DO <-  match("DO", names(Data2))
Temp_C <-  match("Temp_C", names(Data2))
vars <- c(SpC, pH, DO, Temp_C)
SeasonBoxplot <- function(z) {
  for (i in vars) {
   print(ggplot(z, aes_(x = ~Season, y = as.name(names(z)[i]), color = ~Season)) +
    geom_boxplot())
          }
}
SeasonBoxplot(Data2)
```

The width of the boxes in the boxplots above show how much a parameter changed within a season. What is the best way to compare the magnitudes of diel cycles between seasons?

### Average % change by season
Here I compare the magnitude of diel cycles in each season by calculating the % change of each measurement from the seasonal mean. But I want to calculate the % change from the daily mean. I'm not sure how to do that.

```{r Average percent change by season}
percent_change <- function(x) {
  x <- 100*((x - mean(x, na.rm = TRUE))/mean(x, na.rm = TRUE))
}
season_change <- function(y) {transmute(y, datetime, 
                  Temp_C = percent_change(Temp_C),
                  SpC = percent_change(SpC),
                  pH = percent_change(pH),
                  DO = percent_change(DO),
                  Season
                  )}
Winter <- filter(Data, Season == "Winter")
WinterChange <- season_change(Winter)
Spring <- filter(Data, Season == "Spring")
SpringChange <- season_change(Spring)
Summer <- filter(Data, Season == "Summer")
SummerChange <- season_change(Summer)
Fall <- filter(Data, Season == "Fall")
FallChange <- season_change(Fall)
SeasonChange <- bind_rows(WinterChange, SpringChange, SummerChange, FallChange)
describeBy(SeasonChange[textvars], list(SeasonChange$Season))
# SeasonBoxplot(SeasonChange)
```

### Conductivity seasonal diagnostics

```{r conductivity seasonal diagnostics}
plot_seasonal_diagnostics(Data, datetime, SpC, .feature_set = c("hour", "month.lbl"), .interactive = TRUE)
```

### Temp seasonal diagnostics

```{r Temp seasonal diagnostics}
plot_seasonal_diagnostics(Data, datetime, Temp_C, .feature_set = c("hour", "month.lbl"), .interactive = TRUE)
```

### pH seasonal diagnostics

```{r pH seasonal diagnostics}
plot_seasonal_diagnostics(Data, datetime, pH, .feature_set = c("hour", "month.lbl"), .interactive = TRUE)
```

### DO seasonal diagnostics

```{r DO seasonal diagnostics}
plot_seasonal_diagnostics(Data, datetime, DO, .feature_set = c("hour", "month.lbl"), .interactive = TRUE)
```

```{r}
sessionInfo()
save.image("MillCreekSeasonality.RData")
```