---
title: "Mill Creek Hydrology March 2021"
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
---

This notebook analyzes data from two USGS gauging stations in Nashville TN during the flooding that occured in March 2021: Station 03430550 Mill Creek near Nolensville and Station 03431060 Mill Creek at Thompson Lane near Woodbine. These sites will be used for sample collection. Discharge data is not always available for station 03430550 and is of poor quality. 

```{r set-global, include=FALSE}
knitr::opts_chunk$set(error = FALSE, echo=FALSE, message=FALSE, warning = FALSE)
```

```{r import packages, results='hide'}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, leaflet, htmltools)
```

### Map of site locations

```{r Map}
SiteLocations <- tribble(~url, ~latitude, ~longitude, ~DrainageAreaSqMiles,
"Mill creek at Nolensville https://waterdata.usgs.gov/nwis/uv?site_no=03430550",36.00914,-86.70152,40.53,
"Mill creek at Thompson Lane https://waterdata.usgs.gov/nwis/uv?site_no=03431060",36.11752,-86.71905,93.4)
m <- leaflet(SiteLocations) %>% addMarkers(lat = ~latitude, lng = ~longitude, popup = ~htmlEscape(url)) %>% addTiles()
m
```

### ESRI World basemap

```{r ESRI World basemap}
m %>% addProviderTiles(providers$Esri.WorldImagery)
```

### Mill Creek near Thompson Lane

```{r download & format Thompson Lane data}
# The begin and end dates can be changed in the url below
# Discharge not always reported, so use stage height instead
TL <- read_delim("https://nwis.waterdata.usgs.gov/usa/nwis/uv/?cb_00060=on&format=rdb&site_no=03431060&period=&begin_date=2021-03-01&end_date=2021-04-01", "\t", escape_double = FALSE, col_names = TRUE, locale = locale(tz = "America/Chicago"), trim_ws = TRUE, skip = 26)
TL <- TL[-1,-(1:2)]
TL <- rename(TL, Discharge = `131090_00060`)
TL <- select(TL, datetime, Discharge)
TL <- type_convert(TL)
```

#### Plot of discharge over time for flood event. 

```{r Thompson Lane discharge plot}
print(ggplot(TL, aes(x = datetime, y = Discharge)) +
  geom_line())
```

The flood event on March 28 2021 had a massive discharge.

### Mill Creek near Nolensville

```{r download & format Nolensville data}
# Discharge not always reported, so use gage height instead
Nolensville <- read_delim("https://nwis.waterdata.usgs.gov/usa/nwis/uv/?cb_00065=on&cb_00045=on&format=rdb&site_no=03430550&period=&begin_date=2021-03-01&end_date=2021-03-31", "\t", escape_double = FALSE, col_names = TRUE, locale = locale(tz = "America/Chicago"), trim_ws = TRUE, skip = 28)
Nolensville <- Nolensville[-1,-(1:2)]
Nolensville <- rename(Nolensville, GageHeight = `131076_00065`, Precip = `131078_00045`)
Nolensville <- select(Nolensville, datetime, GageHeight, Precip)
Nolensville <- type_convert(Nolensville)
```

Create plots of gage height and precipitation over time for flood event. Discharge is measured every 5 minutes, but precipitation only every 15 minutes.

```{r Nolensville plots}
print(ggplot(data = Nolensville) +
  geom_line(aes(x = datetime, y = GageHeight)) +
  ylab("Gage Height (ft)") +
  theme_bw())
Nolensville <- Nolensville %>% drop_na()
print(ggplot(data = Nolensville) +
  geom_line(aes(x = datetime, y = Precip), color = "blue") +
  ylab("Precipitation (inches)") +
  theme_bw())
```

The peak in gage height occurs shortly after each precipitation event. The high discharge events are short-lived due to rapid runoff in an urban setting with a hig proportion of ground surface that is paved and impermeable. The "flashy" nature is also a reflection of the watershed area upstream from the Nolensville station being only 40.5 square miles.