---
title: "Long-term trends in salinity and pH in Mill Creek"
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
---

This notebook tests whether salinization syndrome has occurred in Mill Creek since 2007. It analyzes temporal trends in conductivity (a measure of salinity in microsiemens per centimeter) and pH measured by the USGS. The site is USGS 03431083 MILL CREEK AT LEBANON ROAD AT NASHVILLE, TN. Information on the site is at <https://waterdata.usgs.gov/nwis/inventory/?site_no=03431083&agency_cd=USGS> Location: Latitude 36°09'50", Longitude 86°41'57" NAD27

#### Reference

Kaushal S. S., Likens G. E., Pace M. L., Utz R. M., Haq S., Gorman J. and Grese M. (2018) Freshwater salinization syndrome on a continental scale. Proc. Natl. Acad. Sci. 115, E574 LP-E583. Available at: <http://www.pnas.org/content/115/4/E574.abstract>

```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE, echo=FALSE, message=FALSE, warning=FALSE)
```

```{r import packages, results='hide'}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, lubridate, skimr, timetk, leaflet, psych, htmltools, mblm)
```

### Map of site location with default Streetmap basemap

```{r Map}
site <- tibble(lat = 36.16389, lng = -86.69917, cap = c("USGS Station 03431083 Mill Creek at Lebanon Road, https://waterdata.usgs.gov/nwis/inventory/?site_no=03431083&agency_cd=USGS"))
m <- leaflet(site) %>% addMarkers(lat = ~lat, lng = ~lng, popup = ~htmlEscape(cap)) %>% addTiles()
m
```

### ESRI World basemap

```{r ESRI World basemap}
m %>% addProviderTiles(providers$Esri.WorldImagery)
```

```{r Read data from USGS website}
# The begin and end dates can be changed in the url below
Data <- read_delim("https://nwis.waterdata.usgs.gov/usa/nwis/uv/?cb_00095=on&cb_00400=on&format=rdb&site_no=03431083&period=&begin_date=2003-09-10&end_date=2021-05-31", "\t", escape_double = FALSE, col_names = TRUE, locale = locale(tz = "America/Chicago"), trim_ws = TRUE, skip = 30)
Data <- Data[-1,-(1:2)]
Data <- rename(Data, SpC = `131094_00095`, pH = `131095_00400`)
Data <- select(Data, datetime, SpC, pH)
Data <- type_convert(Data)
skim(Data)
write_excel_csv(Data, "SalinizationSyndromeMillCreek.csv", col_names = TRUE)
```

### Plots of temporal trends

```{r temporal trends}
DataLong <-  gather(Data, variable, measurement, SpC:pH, factor_key=TRUE)
MinDate <- as.Date(min(Data$datetime))
MaxDate <- as.Date(max(Data$datetime))
print(paste("From ", MinDate, " to ", MaxDate))
print(ggplot(DataLong, aes(x = datetime, y = measurement)) +
  geom_line() +
  scale_x_datetime(date_labels = "%Y-%m") +
  facet_wrap(~ variable, scales = "free", nrow = 2))
```

### Thiel-Sen statistics for SpC

```{r Thiel-Sen statistics SpC}
# Replace records with multiple measurements in the same week with median value
year_week <- function(x,base) week(x) - week(base) + 52*(year(x) - year(base))
Data$Week <- year_week(Data$datetime, Data$datetime[1])
Data2 <- Data %>%
    drop_na() %>% 
    group_by(Week) %>%
    summarize(
      SpC = median(SpC),
      pH = median(pH)
  )
  fit <- mblm(SpC~Week, Data2, repeated = FALSE)
  print(paste("SpC from ", min(Data$datetime), " to ", max(Data$datetime)))
  print(summary(fit))
  print(ggplot(Data2, aes(x = Week, y = SpC)) +
    geom_point() +
      geom_abline(intercept = fit$coefficients[1], slope = fit$coefficients[2]) 
   )
```

SpC actually decreased between 2003 and 2021 (slope = -0.02).

### Thiel-Sen statistics pH

```{r Thiel-Sen statistics pH}
 fit <- mblm(pH~Week, Data2, repeated = FALSE)
print(paste("pH from ", min(Data$datetime), " to ", max(Data$datetime)))  
print(summary(fit))
  print(ggplot(Data2, aes(x = Week, y = pH)) +
    geom_point() +
      geom_abline(intercept = fit$coefficients[1], slope = fit$coefficients[2]) 
   )
```

Slope for pH = 0, so no change over the fourteen-year period.
