---
title: "2022 East Fork Creek Stream Metabolism Summary"
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
---

This notebook is for creating summary plots of stream metabolism parameters estimated from dissolved osygen time series using the BASEmetab package of Grace et al. (2015). In-situ measurements collected from East Fork Creek in Franklin, TN by John C. Ayers, Weizhuo Jing, Kevin Chen, Ella Daugherty, and Gabriel Perez in Summer 2022.  
Reference:
Grace MR, Giling DP, Hladyz S, et al (2015) Fast processing of diel oxygen curves: Estimating stream metabolism with BASE (BAyesian Single-station Estimation). Limnol. Oceanogr. methods / 13:e10011

```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```

### Useful packages

```{r import packages}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, skimr)
```

```{r}
# setwd("C:/Users/ayersj.VANDERBILT/OneDrive - Vanderbilt/Shared Documents/Pilot Project - Water Quality/WaterQualityCode/2022/EFC")
Data <- read_csv("2022EFC_StreamMetabolismSummary.csv", 
    col_types = cols(Date = col_date(format = "%m/%d/%Y"), 
        Device = col_factor(levels = c("MiniDOT", 
            "Exo2"))))
skim(Data)
```

```{r}

DataLong <- Data %>%
  rename(GPP = GPP.mean, ER = ER.mean, NEP = NEP.mean) %>%
  gather(GPP, ER, NEP, key = "Parameter", value = "Value")
  
DataLong$max <- c(0)
DataLong$min <- c(0)
j <- nrow(DataLong)
for (i in 1:j) {
  if (DataLong$Parameter[i] == "GPP") {
  DataLong$max[i] <- DataLong$Value[i] + DataLong$GPP.sd[i]
  DataLong$min[i] <- DataLong$Value[i] - DataLong$GPP.sd[i]
}  
else if (DataLong$Parameter[i] == "ER") {
  DataLong$max[i] <- DataLong$Value[i] + DataLong$ER.sd[i]
  DataLong$min[i] <- DataLong$Value[i] - DataLong$ER.sd[i]
} 
  else {
  DataLong$max[i] <- DataLong$Value[i] + DataLong$NEP.sd[i]
  DataLong$min[i] <- DataLong$Value[i] - DataLong$NEP.sd[i]
}
}
DataLong$Parameter <- factor(DataLong$Parameter)
DataPlot <- DataLong %>%
  select(Date, Device, Parameter, Value, max, min)
 

```

```{r}
p <- ggplot(DataPlot, aes(x = Date, y = Value, color = Parameter)) +
  geom_errorbar(aes(ymin = min, ymax = max)) +
  geom_point(aes(shape = Device)) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()
print(p)
ggsave("2022_EFC_StreamMetab_ErrorBars.png", width = 2000, height = 2*618, units = "px")

```

Try plotting without outliers on 2022-10-25

```{r}
FilteredDataPlot <- filter(DataPlot, Date != "2022-10-25")
p <- ggplot(FilteredDataPlot, aes(x = Date, y = Value, color = Parameter)) +
  geom_errorbar(aes(ymin = min, ymax = max)) +
  geom_point(aes(shape = Device)) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()
print(p)
ggsave("2022_EFC_StreamMetab_ErrorBars_Filtered.png", width = 2000, height = 2*618, units = "px")

```


```{r}

DataMean <- DataPlot %>%
  group_by(Date, Parameter) %>%
  filter(Parameter != "NEP") %>%
  summarize(
       Value = mean(Value)
      )

j <- nrow(DataMean)
for (i in 1:j) {
   if(DataMean$Parameter[i] == "ER") {
    DataMean$Value[i] <- -DataMean$Value[i]
   }
}
 
DataMean

p <- ggplot(DataMean, aes(x=Date, y = Value, fill = Parameter)) +
  geom_area() +
  scale_fill_manual(values = c("brown", "green")) +
  ylim(-35,35) +
  ylab("Value (mg O2 L-1 day-1)")
print(p)
ggsave("2022_EFC_GPPvsER.png", width = 2000, height = 2*618, units = "px")

```

